---
title: "Brite hiearchy as dataframe in R"
author: "Casper Sahl Poulsen"
date: '10082020'
always_allow_html: yes
output:
  pdf_document:
    toc: yes
    toc_depth: '5'
  html_document:
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
Hacky solution to creating a brite hiearchy as dataframe in R.   
The hiearchy is retrieved as json from KEGG orthology  
https://www.genome.jp/kegg-bin/get_htext#C1  

I have not found any better solution to manipulate a nested list and thought this might be helpful for this very specific purpose. I'm sure there is a better solution, but sometimes you just want it to work. 

### Packages 
```{r}
#install.packages(c("plyr", "dplyr", "rjson", "stringr"))
library("plyr")
library("dplyr")
library("rjson")
library("stringr")
```

## Analysis
### Getting the data to dataframe
Raw dataframe
```{r}
# KObrite <- fromJSON(file = "ko00001.json") #Stored locally
KObrite <- fromJSON(file = 
                      "https://www.genome.jp/kegg-bin/download_htext?htext=ko00001&format=json&filedir=")
#print(KObrite)
#lengths(KObrite)

str(KObrite, max.level = 1)
str(KObrite, max.level = 2)
str(KObrite, max.level = 3)
str(KObrite, max.level = 4)
#str(KObrite, max.level = 5)

#Make unlisted version
KObriter<-unlist(KObrite, recursive=TRUE)
KObriter[1:10]

#Got this from Claus Wilke:
#https://stackoverflow.com/questions/48250997/extracting-information-from-multi-level-nested-lists/48251877
KO_flat <- function(x) {
  if (is.null(x)) return(NULL)
  if (!is.null(x$name)) { # if there's a name, we're at the level of a person
    children <- x$children
    x$children <- NULL
    row <- data.frame(plyr::defaults(x, list(name = NA)),
                      stringsAsFactors = FALSE)
    rbind(row, KO_flat(children))
  }
  else {
    purrr::map_dfr(x, KO_flat)
  }
}
flat<-KO_flat(KObrite)
flat[1:10,]

##pseudo hiearchy
KObriteHier<-names(rapply(KObrite, function(x) head(x, 1)))
#which(KObriteHier=="children.name")
#which(KObriteHier=="children.children.name")
#which(KObriteHier=="children.children.children.name")
#length(which(KObriteHier=="children.children.children.children.name"))
#length(which(KObriteHier=="children.children.children.children.name")[indexing[1]:indexing[2]]) #Select first indexing to check

##The loop is build to repeat the top level hierchy based on the number of times "children.children.children.children.name" were observed in the given interval    
##Level 1
#flat[which(KObriteHier=="children.name"),]
vec<-character()
indexing<-c(which(KObriteHier=="children.name"), length(KObriteHier))
#i<-1
p<-0
for (i in which(KObriteHier=="children.name")) {
  p<-p+1
  vec<-c(vec, rep(flat[i,], 
                  times=length(which(
                      KObriteHier[indexing[p]:indexing[p+1]]==
                        "children.children.children.children.name"))))
}
level1<-vec

##Level 2
#flat[which(KObriteHier=="children.children.name"),]
vec<-character()
indexing<-c(which(KObriteHier=="children.children.name"), length(KObriteHier))
#i<-1
p<-0
for (i in which(KObriteHier=="children.children.name")) {
  p<-p+1
  vec<-c(vec, rep(flat[i,], 
                  times=length(which(
                      KObriteHier[indexing[p]:indexing[p+1]]==
                        "children.children.children.children.name"))))
}
level2<-vec

##Level 3
#flat[which(KObriteHier=="children.children.children.name"),]
vec<-character()
indexing<-c(which(KObriteHier=="children.children.children.name"), length(KObriteHier))
#i<-1
p<-0
for (i in which(KObriteHier=="children.children.children.name")) {
  p<-p+1
  vec<-c(vec, rep(flat[i,], 
                  times=length(which(
                      KObriteHier[indexing[p]:indexing[p+1]]==
                        "children.children.children.children.name"))))
}
level3<-vec

##Level 4 
#straight forward at lowest level just need all "children.children.children.children.name"
level4<-flat[which(KObriteHier=="children.children.children.children.name"),]


df<-data.frame(Level1=level1, Level2=level2, Level3=level3, Level4=level4)
```

### Manipulate dataframe
Just some minor manipulations that I used myself
```{r}
#Not interested in level 4 
df2<-df[,1:3] %>% distinct

#Only want as part of a path used as ID
df2$ID <- as.character(regmatches(df2$Level3, gregexpr("\\[.+?\\]", df2$Level3)))
df2$ID <- str_replace(df2$ID, "\\[PATH:","")
df2$ID <- str_replace(df2$ID, "\\]","")

#Remove numbers 
df2$Level1<-gsub("^\\d+|\\d+$", "", df2$Level1)    
df2$Level1<-trimws(df2$Level1)

df2$Level2<-gsub("^\\d+|\\d+$", "", df2$Level2)    
df2$Level2<-trimws(df2$Level2)

df2$Level3<-gsub("^\\d+|\\d+$", "", df2$Level3)    
df2$Level3<-trimws(df2$Level3)

#write.table(df2, file="FeatureFunctional_200812.txt", quote = F, row.names = F, sep="\t")
```


## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r date}
Sys.Date()
```

